
\documentclass[12pt]{amsart}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{algorithm}
\usepackage{pifont}
\usepackage{algpseudocode}
\usepackage{caption}
\usepackage{enumerate}
\usepackage{float}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{tcolorbox}
\tcbuselibrary{skins, breakable}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{listings}% http://ctan.org/pkg/listings
\lstset{
	basicstyle=\ttfamily,
	language=Python,
	showstringspaces=false,
	mathescape
}
\usepackage{blkarray}
\newcommand{\matindex}[1]{\mbox{\scriptsize#1}}% Matrix index

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{question}[theorem]{Question}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\cF}{\mathcal F}
\newcommand{\cD}{\mathcal D}
\newcommand{\cP}{\mathcal P}
\newcommand{\cR}{\mathcal R}
\newcommand{\cS}{\mathcal S}
\newcommand{\fS}{\mathfrak S}
\newcommand{\R}{\mathbb{R}}
\newcommand{\dist}{\text{dist}}
\newcommand{\pf}{\text{pf}}
\newcommand{\PP}{\text{P}}
\newcommand{\NP}{\text{NP}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\ord}{\text{ord}}
\newcommand{\OPT}{\text{OPT}}
\newcommand{\cost}{\text{cost}}
\newcommand{\price}{\text{price}}
\newcommand{\E}{\text{E}}

\tcbset{breakable, enhanced jigsaw, parbox=false, title=Solution.}

% Header ---------------------------------------------------------------------------
\lhead{CS 8803GA}
\chead{Homework 1 Solutions}
\rhead{\today}
% ----------------------------------------------------------------------------------

\begin{document}
	\thispagestyle{fancy}
	\pagestyle{plain}
	
	\noindent{\bf Problem 0.1:}  \\
	(a)  [DPV] 0.1(c): $f(n) = 100n + \log n, g(n) = n + (\log n)^2$
	
	\begin{tcolorbox}
		C: $f(n) = O(g(n))$ and $g(n) = O(f(n))$\\
		$\log n$ and $(\log n)^2$ are both $O(n)$, and can be ignored when comparing $f(n)$ and $g(n)$.  Ignoring constant factors, both $f(n)$ and $g(n)$ are $\Theta(n)$, so $f(n) = O(g(n))$ and $g(n) = O(f(n))$.
	\end{tcolorbox}
	
	(b)  [DPV] 0.1(d): $f(n) = n \log n, g(n) = 10n \log 10n$
	
	\begin{tcolorbox}
		C: $f(n) = O(g(n))$ and $g(n) = O(f(n))$\\
		$\log 10n = \log 10 + \log n$, and $\log 10$ can be ignored here.  Ignoring constant factors, both $f(n)$ and $g(n)$ are $\Theta(n\log n)$, so $f(n) = O(g(n))$ and $g(n) = O(f(n))$.
	\end{tcolorbox}
	
	(c)  [DPV] 0.1(k): $f(n) = \sqrt{n}, g(n) = (\log n)^3$
	
	\begin{tcolorbox}
		B: $g(n) = O(f(n))$\\
		Taking the cube root of each $f(n)$ and $g(n)$ yields $n^{1/6}$ and $\log n$.  Any positive power of $n$, including powers less than 1, will dominate $\log n$, so $g(n) = O(f(n))$.
	\end{tcolorbox}

	(d) [DPV] 0.1($\ell$): $f(n) = \sqrt{n}, g(n) = 5^{\log_2 n}$
	
	\begin{tcolorbox}
		A: $f(n) = O(g(n))$\\
		\[  5^{\log_2 n} = \left(2^{\log_2{5}}\right)^{\log_2 n} = \left(2^{\log_2{n}}\right)^{\log_2{5}} = n^{\log_2{5}}.\]
		\\
		Since $\log_2{5} > 1$ we have that $n^{\log_5{2}} > n >  n^{.5}$. Therefore, $f(n) = O(g(n))$.
	\end{tcolorbox}

	\newpage
	\noindent{\bf Problem 6.2: Hotel stops with minimum penalty }  \\
	(a) Define the entries of your table in words.  E.g., $T(i)$ or $T(i,j)$ is ...
	
	\begin{tcolorbox}
		$T(i)$ is the minimum penalty obtainable for the trip from mile 0 to mile $a_i$, with the last stop at hotel $i$.
	\end{tcolorbox}

	(b) State recurrence for entries of table in terms of smaller subproblems.

	\begin{tcolorbox}
		Each entry $T(i)$ is computed as the minimum over all previous hotels $k$ of the minimum penalty to get to hotel $k$ plus the penalty from hotel $k$ to hotel $i$.\\
		$T(i) = \min_k\{T(k) + (200 - (a_i - a_k))^2 : 0 \leq k \leq i-1\}$
	\end{tcolorbox}

	(c) Write pseudocode for your algorithm to solve this problem.
	
	\begin{tcolorbox}
		\begin{lstlisting}[tabsize=4]
$T(0) = 0$
for $i=1$ to $n$:
	$T(i) = (200 - a_i)^2$
	$prev(i) = NULL$
	for $k=1$ to $i-1$:
		if $T(i) > T(k) + (200 - (a_i - a_k))^2$ then
			$T(i) = T(k) + (200 - (a_i - a_k))^2$
			$prev(i) = k$

# $T(n)$ is now the minimum penalty
# Next, output the optimal sequence of hotels
$i = n$
$output(i)$
while $prev(i) \neq NULL$:
	$i = prev(i)$
	$output(i)$

		\end{lstlisting}
	\end{tcolorbox}

	(d) Analyze the running time of your algorithm.
	
	\begin{tcolorbox}
		Each entry in $T(i)$ takes $O(n)$ time to compute, since each is the minimum over up to $n-1$ expressions.  There are $n$ entries in $T$, so this algorithm takes $O(n^2)$ time.
	\end{tcolorbox}
	
	\newpage
	\noindent{\bf Problem 6.3: Yuckdonald's }  \\
	(a) Define the entries of your table in words.  E.g., $T(i)$ or $T(i,j)$ is ...
	
	\begin{tcolorbox}
		$T(i)$ is the maximum profit from a valid subset of locations from $m_1, m_2, ... m_i$ that includes $m_i$.
	\end{tcolorbox}
	
	(b) State recurrence for entries of table in terms of smaller subproblems.
	
	\begin{tcolorbox}
		Each entry $T(i)$ is computed as the profit from opening location $i$, plus the maximum profit from previous valid sets of locations at least $k$ miles away from location $i$.\\
		$T(i) = p_i + \max_j\{T(j) : j < i, m_j \leq m_i - k\}$
	\end{tcolorbox}
	
	(c) Write pseudocode for your algorithm to solve this problem.
	
	\begin{tcolorbox}
		\begin{lstlisting}[tabsize=4]
$T(0) = 0$
for $i=1$ to $n$:
	$T(i) = p_i$
	for $j=1$ to $i-1$:
		if $m_j \leq m_i - k$ then
			if $T(i) < T(j) + p_i$ then
				$T(i) = T(j) + p_i$

return $\max_i L(i)$
		\end{lstlisting}
	\end{tcolorbox}
	
	(d) Analyze the running time of your algorithm.
	
	\begin{tcolorbox}
		Each entry in $T(i)$ takes $O(n)$ time to compute, since each is the maximum over up to $n-1$ expressions.  There are $n$ entries in $T$, so this algorithm takes $O(n^2)$ time.
	\end{tcolorbox}
	
\end{document}







